DeepSphere: a graph-based spherical CNN

### 2019-09-19 main.tex
Updates from Overleaf

\cite{khasanova2017graphomni} designed a discrete Laplacian that is explicitly intended to work on the sphere with the equiangular sampling. They consider the set $\mathcal G$ of all the possible graphs where each node is connected only to four of its nearest neighbours (North, Sud, West, East) and propose a weighting scheme $w_{ij}$ that minimizes the difference in the response to the polynomial spectral filter $\mathcal F = \mathbf L$ evaluated on images of the same object seen at different latitudes. In other words, they solve the minimization problem
\begin{equation} \label{eq:minimization frossard}
	\min_{W\in\mathcal G} \left|\mathcal{F}\left(\mathbf{y}\left(v_{ e}\right)\right)-\mathcal{F}\left(\mathbf{y}\left(v_{ i}\right)\right)\right|
\end{equation}
for the adjacency matrix $W$, where $\mathbf y(v_i)$ is the image of the object on the sphere centered on the vertex $v_i$, and $\mathcal F (\mathbf y(v_e))$ is the response of the filter at the vertex $v_e$ that lies at the same longitude of the vertex $v_i$ but on the equator. In their work they prove that the optimal weights solving the minimization problem (\ref{eq:minimization frossard}) are given by weights $w_{ij}$ inversely proportional to the Euclidean distance between vertices:
\begin{equation} \label{eq:frossard weights}
	w_{ij} = \frac{1}{\norm{x_i-x_j}}
\end{equation}

@@

@@

% The plan for this section
% 0) Explain why the Laplacian is so important...
By definition, the graph convolution (on the sphere) is rotation equivariant to rotation, if and only if, it commutes with the rotation operator. 
As, in the context of this work, graph convolution can be performed by successive application of the graph Laplacian, i.e: $g(\b{L}) \b{f} = \left(\sum_{i=0}^K \alpha_i \b{L}^i\right) \b{f}$, if the graph Laplacian $\b{L}$ commutes with the rotation operator, then, by recursion, it will also commute with the convolution $g(\b{L})$.
Hence to simplify our analysis, we concentrate on the graph Laplacian.

% 0.5) Explain the sampling and metric
We first need to deal with the sampling problem. 
Take a sampling scheme $V=\{v_i\in\mathbb S^2, i=0, \dots, n\}$ of the sphere, a weighted undirected graph $G(V, E, \mathbf W)$, a signal $f: \mathbb S^2\to\mathbb R$ and its sampled representation $\mathbf f:\ f_i=f(v_i)$. Suppose that there exists a sampling operator $T_V: L^2(\mathbb S^2) \supset F\to \mathbb R^n,\  T_V(f) = \mathbf f$ defined on a suitable subspace $F$ of $L^2(\mathbb S^2)$ such that it is invertible, i.e., we can unambiguously reconstruct the function $f\in F$ from its sampled values $\mathbf f$. The existence of such subspace depends on the sampling scheme $V$, and its characterization is a common problem in signal processing \cite{Driscoll:1994:CFT:184069.184073}. Define the rotation operator $R(g), g\in SO(3)$: $R(g) f(\omega) = f\left(g^{-1} \omega\right)$.

We now want to understand how to set the edges and the weights of $G$ such that
\begin{equation} \label{eq:equivariance}
	T \R(g) T^{-1} \Omega_k T f = \Omega_k T R(g) f,
\end{equation}
i.e., the graph convolution $\Omega_k$ and any rotation $R(g)$ commute.

Verifying equation (\ref{eq:equivariance}) is really hard in practice, due to the fact that for almost all sampling schemes $V$ it is not known if there exists a space $F$ in which $T$ is invertible. A special case is the \textit{equiangular sampling} scheme where a sampling theorem holds \cite{Driscoll:1994:CFT:184069.184073}, and thus we have a closed form of $T^{-1}$. For sampling schemes where no such sampling theorem is available, there are implementations of discrete SHT to reconstruct a sampled signal $\mathbf f$, thus providing a way to approximate $T^{-1}$. Thanks to this we are able, for a given sampling, a given function $f$, a given rotation $g$, and a given kernel $k$, to compute the \textit{normalized equivariance error}
\begin{equation} \label{eq:equivariance error}
	E_{G}(f, g) = \left(\frac{ \norm {T R(g) T^{-1} \Omega_k Tf - \Omega_k T R(g) f}_{L^2(\mathbb R^2)}}{\norm {Tf}_{L^2(\mathbb R^2)}}\right)^2,
\end{equation}
where $T^{-1}$ is substituted with a discrete SHT in case $T$ is not invertible.
A measure of how equivariant a graph is with respect to rotations will then be given by the \textit{mean equivariance error}
\begin{equation} \label{eq:mean equivariance error}
	\overline E_G = \mathbb E_{f, g} \ E_G(f, g).
\end{equation}
In practice the expected value is obtained by averaging over a finite number of random functions and random rotations. The mean equivariance error $\overline E_G$ gives us an indication of how close the graph $G$ is from being equivariant to rotations. 
Of course result will depend on the sampling and on the graph construction.

% !) summary the work of Khasanova -> No! to be put in section 2) Method

% 2) explain our approach
We inspire ourselves with Belkin, which says that for \emph{uniform sampling} the graph Laplacian converges toward the Laplace-Beltrami. b) The Laplace Beltrami commute with rotation. The following theorem extend the work of Belkin and Nyiogi to fixed sampling \nati{(Including Healpix or not?)}. 
Given a sampling $V = \{x_0, \dots, x_{n-1}\}$ define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, define $A_i$ to be its corresponding area and $d_i$ to be the radius of the smallest ball in $\mathbb R^3$ containing $\sigma_i$. Define $d^{(n)} := \max_{i=0, \dots, n}d_i$ and $A^{(n)}=\max_{i=0, \dots, n}A_i$.

@@

This is \emph{not} a proof of equivariance of the graph convolution for a specific sampling. We simply observe that the graph Laplacian converges pointwisely toward an operator that commutes with rotation. To have the desired equality in eq..., we would need to prove a stronger result, for example uniform convergence (in the space of the operators).
\nati{Should we put this here?: }This does not hold in general \cite{belkin-spectral}. However, we believe that this is achievable for a bandlimited signals.
More importantly, the proof of theorem \ref{theo:pointwise convergence for a regular sampling} in Appendix ... inspires us in the construction of our graph Laplacian. a) use gaussian kernel b) we need to adjust t with respect of n such that ...

% 3) Show experimental results
Define the error, show results and comments
\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{figure_architecture_v3}
	\caption{Example architecture from \cite{perraudin2019deepsphere}.
		Global tasks need a spatial summarization: the FCN variant is rotation invariant (and accepts inputs of varying sizes), while the CNN variant is not.
		Dense tasks (when the output lives on the sphere, like segmentation) are rotation equivariant.
	}
	\label{fig:architecture}
\end{figure}

@@

>>>>

% \cite{khasanova2017graphomni} designed a discrete Laplacian that is explicitly intended to work on the sphere with the equiangular sampling. They consider the set $\mathcal G$ of all the possible graphs where each node is connected only to four of its nearest neighbours (North, Sud, West, East) and propose a weighting scheme $w_{ij}$ that minimizes the difference in the response to the polynomial spectral filter $\mathcal F = \mathbf L$ evaluated on images of the same object seen at different latitudes. In other words, they solve the minimization problem
% \begin{equation} \label{eq:minimization frossard}
% 	\min_{W\in\mathcal G} \left|\mathcal{F}\left(\mathbf{y}\left(v_{ e}\right)\right)-\mathcal{F}\left(\mathbf{y}\left(v_{ i}\right)\right)\right|
% \end{equation}
% for the adjacency matrix $W$, where $\mathbf y(v_i)$ is the image of the object on the sphere centered on the vertex $v_i$, and $\mathcal F (\mathbf y(v_e))$ is the response of the filter at the vertex $v_e$ that lies at the same longitude of the vertex $v_i$ but on the equator. In their work they prove that the optimal weights solving the minimization problem (\ref{eq:minimization frossard}) are given by weights $w_{ij}$ inversely proportional to the Euclidean distance between vertices:
% \begin{equation} \label{eq:frossard weights}
% 	w_{ij} = \frac{1}{\norm{x_i-x_j}}
% \end{equation}

@@

% \section{Notation}
% Throughout this contribution, we denote with bold font all vectors and matrices such as $\b{L}$ the graph Laplacian and $\b{f}$ the vector of the sampled function $f$.
% \nati{Right now, it is not True}
% % , a weighted undirected graph $G(V, E, \mathbf W)$,
% \nati{I do not think we should have a notation section, but should somehow write this somewhere...}

@@

% The plan for this section
% 0.1 Explain the sampling problem
\paragraph{Sampling} 
As the graph convolution act on a sampled version of the signal $\b{f}$ and the rotation on its continuous counterpart $f$, we first need to define a sampling operator.
Given a sampling scheme $V=\{v_i\in\mathbb S^2, i=0, \dots, n\}$ of the sphere and a function $f : \mathbb S^2 \supset F_V \to \mathbb R$, the sampling operator $T_V: L^2(\mathbb S^2) \supset F\to \mathbb R^n,\  T_V(f) = \b{f}$ is defined as $\b{f}:\ f_i=f(v_i)$, where $F_V$ suitable subspace of $L^2(\mathbb S^2)$ such that $T_V$ is invertible, i.e., we can unambiguously reconstruct the function $f\in F_V$ from its sampled values $\b{f}$.
The existence of such subspace depends on the sampling scheme $V$, and its characterization is a common problem in signal processing \cite{Driscoll:1994:CFT:184069.184073}. 
For most samplings, it is not known if $F_V$ exist and hence if $T_V$ is invertible. A special case is the \textit{equiangular sampling} scheme where a sampling theorem holds, and thus we have a closed form of $T_V^{-1}$ is known. % \cite{Driscoll:1994:CFT:184069.184073}
For sampling schemes where no such sampling formula is available, we leverage the discrete SHT to reconstruct $f$ from $\bf{f}$, thus approximating $T_V^{-1}$. 
For all theoretical consideration, we assume that $F_V$ exists and $f \in F_V$.

% 0.2 Explain why the Laplacian is so important...
\paragraph{Problem formulation} 
By definition, our (spherical) graph convolution is rotation equivariant, if and only if, it commutes with the rotation operator defined as $R(g), g\in SO(3)$: $R(g) f(x) = f\left(g^{-1} x \right)$. 
In the context of this work, graph convolution is performed by recursive applications of the graph Laplacian, i.e: $h(\b{L}) \b{f} = \left(\sum_{i=0}^K \alpha_i \b{L}^i\right) \b{f}$\nati{Put this equation somewhere else?}.  Hence, if $\b{L}$ commutes with the rotation operator, then, by recursion, it will also commute with the convolution $h(\b{L})$.
As a result, to simplify our analysis, we concentrate on the graph Laplacian and $h(\b{L})$ is rotational invariant if and only if 
% 0.3 Formulate the problem
\begin{equation} \label{eq:equivariance}
	\b{R}_V(g) \b{L} \b{f} = \b{L} \b{R}_V(g) \b{f} \hspace{1cm} \forall f\in F_V \text{ and } \forall g\in SO(3), 
\end{equation}
where $\b{R}_V(g) = T_V R(g) T_V^{-1}$. For practical application, we define the \textit{normalized equivariance error} for a signal $\b{f}$, and a  rotation $g$ as:
\begin{equation} \label{eq:equivariance error}
	E_{\b{L}}(\b{f}, g) = \left(\frac{ \norm {\b{R}_V(g) \b{L} \b{f} - \b{L} \b{R}_V(g) \b{f}} }{\norm {\b{f}}}\right)^2,
\end{equation}
More generally for a class of signal $f \in C \subset F_V$, the \textit{mean equivariance error}
\begin{equation} \label{eq:mean equivariance error}
	\overline E_{\b{L}, C} = \mathbb E_{\b{f}\in C, g\in SO(3)} \ E_{\b{L}}(\b{f}, g).
\end{equation}
In practice, to anlyse the equivariance error frequency,  the set $C$ we used is composed of functions $f$ with made of spherical harmonic of a single degree $\ell$.
The expected value is obtained by averaging over a finite number of random functions and random rotations.

\subsection{Finding a the optimal set of weights}
% !) summary the work of Khasanova -> No! to be put in section 2) Method
\cite{khasanova2017graphomni} designed a Laplacian $\b{L}$ explicitly intended to work with the equiangular sampling. 
They consider the set $\mathcal{G}$ of all the possible graphs where each node is connected only to four of its nearest neighbours (North, Sud, West, East) and propose a weighting scheme $w_{ij}$ that minimizes \eqref{eq:equivariance error} for specific rotations on the longitude latitude axis.
Their approach leads to weights $w_{ij}$ inversely proportional to the Euclidean distance between vertices:
\begin{equation} \label{eq:frossard weights}
	w_{ij} = \frac{1}{\norm{x_i-x_j}}
\end{equation}

% 2) explain our approach
We chose a different approach and take inspiration from \cite{belkin2005towards}, which proves that for a random \emph{uniform sampling} scheme the graph Laplacian converges toward the Laplace-Beltrami operator $\Delta$ as the number of sampling point goes to infinity.
This convergence result is important as a) $\Delta$ commutes with rotation and b) is diagonalized by the spherical harmonics. 
In this case the weighting scheme is full (every node is connected to every node) and is based the exponential kernel: 
\nati{@martino, please check this equation... Do you use 4t?}
\begin{equation} \label{eq:belkin_weights}
w_{ij} = e^{-\frac{1}{t} \norm{x_i-x_j}^2}
\end{equation}
In practice
The following theorem extend the work of Belkin and Nyiogi to fixed sampling \nati{(Including Healpix or not?)}. 
Given a sampling $V = \{x_0, \dots, x_{n-1}\}$ define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, define $A_i$ to be its corresponding area and $d_i$ to be the radius of the smallest ball in $\mathbb R^3$ containing $\sigma_i$. Define $d^{(n)} := \max_{i=0, \dots, n}d_i$ and $A^{(n)}=\max_{i=0, \dots, n}A_i$.

@@

This is \emph{not} a proof of equivariance of the graph convolution for a specific sampling. We simply observe that the graph Laplacian converges pointwisely toward an operator that commutes with rotation. To have the desired equality in eq..., we would need to prove a stronger result, for example uniform convergence (in the space of the operators).
\nati{Should we put this here?: }
This does not hold in general \cite{belkin2007convergence}. However, we believe that this is achievable for a bandlimited signals.
More importantly, the proof of theorem \ref{theo:pointwise convergence for a regular sampling} in Appendix ... inspires us in the construction of our graph Laplacian. a) use gaussian kernel b) we need to adjust t with respect of n such that ...

% 3) Show experimental results
Define the error, show results and comments

@@

% \begin{table}[]
%     \centering
%     \begin{tabular}{c|c}
%          &  \\
%          & 
%     \end{tabular}
%     \caption{Caption}
%     \label{tab:cosmo_newgraph}
% \end{table}
